---
title: "p8105_hw3_yq2279"
author: "Qi Yuchen"
date: "2019/10/10"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(patchwork)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 12, 
  fig.asp = .618,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

# Problem 1

### A short description of the dataset

```{r}
library(p8105.datasets)
data("instacart")

# Display the structure of the data
str(instacart)
```

The size of the dataset is `r dim(instacart)[1]` observations and `r dim(instacart)[2]` variables.   
The data structures in this dataset are integer and character.  
The key variables include "reordered" which indicates if this prodcut has been ordered by this user before, "aisle" which is the name of the aisle, "department" which is the name of the department, and "product_name" which is the name of the product.  
To illustrate the observations, take the first observation as an example. The user with user_id 112108 buys the product Bulgarian Yogurt which has been bought by this costumer before indicated by the variable reordered. This product belong to the yogurt	aisle in the dairy eggs department. Put the first to eighth observations together, they form one order from user 112108, which is this user's fourth order and placed at 10 at the day. It has been 9 days since the last order. 

### Number of aisles and which aisles the most items are ordered from

```{r}
# number of aisles
summarise(instacart, n_aisles = n_distinct(aisle_id)) %>% 
  knitr::kable()

# aisles that the most items are ordered from 
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs_aisle = n()) %>% 
  filter(min_rank(desc(n_obs_aisle)) < 4) %>% # here we show the top 3
  arrange(desc(n_obs_aisle)) %>% 
  knitr::kable()
```

There are 134 aisles, and the aisle that the most items are ordered from is fresh vegetables, followed by fresh fruits and packaged vegetables fruits.

### A plot that shows the number of items ordered in each aisle with more than 10000 items ordered

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs_aisle = n()) %>% 
  filter(n_obs_aisle > 10000) %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n_obs_aisle)) %>% 
  ggplot(aes(x = aisle, y = n_obs_aisle, fill = aisle)) + 
  geom_bar(stat = "identity") +
  labs(
    title = "Number of items in each aisle",
    x = "aisle",
    y = "number of items"
  ) +
  coord_flip()

```

### A table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”

```{r}
instacart %>%
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_product = n()) %>% 
  filter(min_rank(desc(n_product)) < 4) %>% 
  arrange(aisle, desc(n_product)) %>% 
  knitr::kable()
```

### A table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered

```{r}
instacart %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  rename(Sunday = "0", Monday = "1", Tuesday = "2", Wednesday = "3", Thursday = "4", Friday = "5", Saturday = "6") %>% 
  knitr::kable(digits = 1)
```

# Problem 2

### Do some data cleaning
```{r}
data("BRFSS")

df_BRFSS = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(location_abbr = locationabbr, location_desc = locationdesc) %>% 
  filter(topic == "Overall Health") %>% 
  filter(
    response == "Excellent" | 
    response == "Very good" | 
    response == "Good" | 
    response == "Fair" | 
    response == "Poor") %>%
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent")))
  
```

### States observed at 7 or more locations in 2002 and 2010

```{r}
# 2002
df_2002 = df_BRFSS %>% 
  filter(year == 2002) %>% 
  group_by(location_abbr) %>% 
  summarise(n_locations = n_distinct(location_desc)) %>% 
  filter(n_locations > 6) 
pull(df_2002, location_abbr)

# 2010
df_2010 = df_BRFSS %>% 
  filter(year == 2010) %>% 
  group_by(location_abbr) %>% 
  summarise(n_locations = n_distinct(location_desc)) %>% 
  filter(n_locations > 6) 
pull(df_2010, location_abbr)
```

In 2002, `r pull(df_2002, location_abbr)` were observed at 7 or more locations. In 2010, `r pull(df_2010, location_abbr)` were observed at 7 or more locations.

### Construct a dataset and make a “spaghetti” plot

```{r}
# limited to Excellent responses, and contains year, state, and a variable that averages the data_value across locations within a state
df_data_mean = df_BRFSS %>% 
  filter(response == "Excellent") %>% 
  rename(state = location_abbr) %>% 
  group_by(state, year) %>% 
  summarize(data_value_mean = mean(data_value, na.rm = TRUE))

# a “spaghetti” plot of this average value over time within a state
df_data_mean %>% 
  ggplot(aes(x = year, y = data_value_mean, color = state)) +
  geom_line() + geom_point() +
  labs(
    title = "Average data value over time within a state",
    y = "average data value"
  ) 
```

### Make a two-panel plot showing distribution of data_value for responses

```{r}
# for the years 2006, and 2010, among locations in NY State
df_BRFSS %>% 
  filter((year == 2006 | year == 2010) & 
         location_abbr == "NY") %>% 
  ggplot(aes(x = response, y = data_value)) + 
  geom_bar(aes(fill = location_desc), stat = "identity", position = "dodge") + 
  labs(title = "Distribution of data_value for responses among locations in NY State") +
  facet_grid(~year)
```

# Problem 3

### Load, tidy, and otherwise wrangle the data

```{r}
df_accel = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(weekday_or_weekend = case_when(
    day %in% c("Friday", "Monday", "Thursday", "Tuesday", "Wednesday") ~ "weekday",
    day %in% c("Saturday", "Sunday") ~ "weekend",
    TRUE ~ ""
  )) %>% 
  select(week, day_id, day,weekday_or_weekend, everything()) %>% 
  mutate(
    day = as.factor(day), 
    week = as.factor(week), 
    weekday_or_weekend = as.factor(weekday_or_weekend)
    ) %>% 
  mutate(day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>% 
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "minite", 
    names_prefix = "activity_", 
    values_to = "minite_counts"
    ) %>% 
  mutate(minite = as.numeric(minite))
```

The variables in this dataset are `r names(df_accel)`, and there are `r dim(df_accel)[1]` observations. The dataset covers 5 weeks which is 35 days.

### Create a total activity variable for each day, and create a table showing these totals

```{r}
n_activity_day = df_accel %>% 
  group_by(week, day) %>% 
  summarize(day_counts = sum(minite_counts)) 

pivot_wider(
  n_activity_day,
  names_from = day,
  values_from = day_counts
  ) %>% 
  knitr::kable(digits = 0)

n_activity_day %>% 
  ggplot(aes(x = day, y = day_counts, color = week, group = week)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Total activity over the day", 
    y = "counts"
    )
```

On Tuesday and Wednsday, activity of different weeks are more stable compared to other days. 

### Make a single-panel plot that shows the 24-hour activity time courses for each day

```{r}
#  use color to indicate day of the week
df_accel %>% 
  ggplot(aes(x = minite, y = minite_counts, color = day, group = day)) +
  geom_line(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  scale_x_continuous(breaks = seq(0, 1400, by = 100), limits = c(0, 1450), labels = seq(0, 1400, by = 100)) +
  labs(
    x = "minute",
    y = "activity counts",
    title = "24-hour activity time courses"
  )
```

Activity counts are peak when it is around 11 am and 9 pm of the day most days.